---
title: "RR2"
author: "Hang Li"
date: "2/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(reshape2)
library(plyr)
library(lattice)
library(lubridate)
```

## Loading data

Download and unzip the file into our WD.

```{r load}
setwd("C:/Users/Ch2ya/Desktop/RR1")
activity <- read.csv("activity.csv")
str(activity)
```
We see that our data is ready for analysis.

## Mean total steps per day

We remove entries without step data and group them by dates. The histogram, mean, and median is given by

```{r mean}
act <- melt(activity[which(!is.na(activity$steps)), ], id.vars = c("date", "interval"))
ss <- dcast(act, date ~ variable, sum)

hist(ss$steps, main = "Histogram of steps per day", xlab = "steps per day", ylab = "days", breaks = 10, col = "purple")
abline(v = mean(ss$steps), lty = 1, lwd = 2)
legend(legend="mean","topright",lty = 1, lwd = 2)

summary(ss$steps)
```


## Daily activity pattern

We now make a time series plot of the average steps taken

```{r time}
stepsint <- dcast(act, interval ~ variable, mean, na.rm = TRUE)
plot(stepsint$interval, stepsint$steps, ty = "l", xlab = "interval", ylab = "steps", main = "Average steps over days on time interval")
```

The time interval of the maximum number of steps is given by

```{r max steps}
stepsint$interval[which.max(stepsint$steps)]
```

4. Imputing missing values
--------------------------

The number of entries with NA as steps is given by 
```{r NA}
nrow(activity[is.na(activity$steps),])
```

One way to impute the missing values is to substitute in the average interval based on the day of the week.
```{r impute }
inputed <- lapply(split(activity, activity$interval), function(x) {
        x$steps[which(is.na(x$steps))] <- mean(x$steps, na.rm = TRUE)
        return(x)})
inputed <- do.call("rbind", inputed)
row.names(inputed) <- NULL
inputed <- split(inputed, inputed$date)
df <- lapply(inputed, function(x) {
        x$steps[which(is.na(x$steps))] <- mean(x$steps, na.rm = TRUE)
        return(x)})
inputed <- do.call("rbind", inputed)
row.names(inputed) <- NULL
act2 <- melt(inputed, id.vars = c("date", "interval"))
steps_sum <- dcast(act2, date ~ variable, sum, na.rm = TRUE)
```

We now compare the histograms
```{r total step per day histogram2}
par(mfrow=c(1,2))
hist(steps_sum$steps, main = "Steps per Day with Imputed Values", 
     xlab = "steps per day", ylab = "days", 
     breaks = 10, col = "red")
abline(v = mean(steps_sum$steps), lty = 1, lwd = 2, col = "black")
legend(x = "topright", "Mean", lty = c(2, 1), lwd = c(2, 2))
hist(ss$steps, main = "Histogram of steps per day", xlab = "steps per day", ylab = "days", breaks = 10, col = "purple")
abline(v = mean(ss$steps), lty = 1, lwd = 2)
legend(legend="mean","topright",lty = 1, lwd = 2)
```

We can we the histograms are slightly different.

## Weekdays vs Weekends

We now lable days with weekdays and weekends

```{r weekday status}
weekdays <- which(weekdays(as.Date(inputed$date)) != "Saturday" &
              weekdays(as.Date(inputed$date)) != "Sunday")
weekends <- which(weekdays(as.Date(inputed$date)) == "Saturday" |
              weekdays(as.Date(inputed$date)) == "Sunday")
temp <- c(rep("a", length(inputed)))
temp[weekends] <- "weekend"
temp[weekdays] <- "weekday"
names(temp) <- "day"
inputed <- cbind(inputed, temp)
names(inputed)[4] <- "day"
inputedsplit <- split(inputed, inputed$day)
weekint <- lapply(inputedsplit, function(x) {
        temp <- aggregate(x$steps, list(x$interval), mean)
        names(temp) <- c("interval", "steps")
        return(temp)
})
weekint <- do.call("rbind", weekint)
weekdays <- grep("weekday" ,row.names(weekint))
weekends <- grep("weekend" ,row.names(weekint))
temp <- c(rep("a", length(weekint$steps)))
temp[weekdays] <- "weekdays"
temp[weekends] <- "weekends"
weekint <- cbind(weekint, temp)
row.names(weekint) <- NULL
names(weekint)[3] <- "day"

ggplot(weekint, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) 
```

We see that step trends varies between weekdays and weekends.

